---
title: "touRnamentofchampions"

output: github_document
---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE
)
```

## 1. Introduction to touRnamentofchampions

touRnamentofchampions is a collection of data sets detailing events across all seasons of Tournament of Champions. It includes Chef information, randomizer information, challenge results, and judge names.

## 2. Installation

Not yet on CRAN. So please use: *devtools::install.packages("celevitz/touRnamentofchampions")*. If it's not appearing to be updated, restart your R sessions, install it again, and call it into your library.

```{r InstallPackage , message=FALSE,warning=FALSE}
devtools::install_github("celevitz/touRnamentofchampions")
```

## 3. News

Issues to work on:

-   Data entry for Seasons 3 and 5
-   The Randomizer data set has duplicate rows for the finals, one for each coast.
-   Complete handedness
-   Add gender to seed dataset

## 4. References & Acknowlegements

Data were collected manually while watching each season of Tournament of Champions.

Huge thanks to <https://github.com/doehm> for all his support!

```{r Library set up, message=FALSE, warning=FALSE,echo=FALSE,eval=TRUE}
library(touRnamentofchampions); library(tidyverse)
```

## 5. Datasets

Across datasets, key joining variables include:

-   `chef`
-   `season`
-   `coast`
-   `region`
-   `round`
-   `episode`

### 5a. Seeds

-   `chef`: Chef name (full name)
-   `season`: Season number
-   `seed`: Seed within their section of the bracket: values of 1 through 8. Chefs that played in the play-in but didn't make the final bracket will have letters in their seeds.
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
-   `nickname`: Guy Fieri's nickname for the chef
-   `handedness`: Whether the chef is righthanded, lefthanded, or ambidextrous

```{r Seeds_table , eval=TRUE}
seeds 
```

### 5b. Randomizer

-   `season`: Season number
-   `episode`: Episode number
-   `round`: Stage of the tournament: Play-in, Round of 32, Round of 16, Quarterfinals, Semifinals, Final
-   `challenge`: Variable to help distinguish rounds within the same Coast & Round
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
-   `randomizer1`: First wheel of randomizer
-   `randomizer2`: Second wheel of randomizer
-   `randomizer3`: Third wheel of randomizer
-   `randomizer4`: Fourth wheel of randomizer
-   `time`: Length of challenge. Unit is minutes
-   `randomizer5`: Fifth wheel of randomizer

```{r Randomizer_table , eval=TRUE}
randomizer 
```

### 5c. Results

-   `season`: Season number
-   `episode`: Episode number
-   `round`: Stage of the tournament: Play-in, Round of 32, Round of 16, Quarterfinals, Semifinals, Final}
-   `challenge`: Variable to help distinguish rounds within the same Coast & Round
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
-   `y`: Numeric value to help when creating the bracket
-   `chef`: Name of chef
-   `commentator`: Who presented their food to the judges: Simon Majumdar or Justin Warner?
-   `order`: When did their food get presented to the judges: Presented 1st or Presented 2nd
-   `score_taste`: Score that chef received for the taste of their dish: values of 0- 50
-   `score_randomizer`: Score that chef received for how well they used the Randomizer: values of 0- 30
-   `score_presentation`: Score that chef received for the presentation of their dish: values of 0- 20
-   `total`: Total score that chef received: between 0 and 100

```{r Results_table , eval=TRUE}
results 
```

### 5d. Judges

The unique identifier is `season`-`episode`, even though a judge may judge more than one round in an episode.

-   `season`: Season number
-   `episode`: Episode number
-   `judge`: Name of guest judge

```{r Judges_table , eval=TRUE}
judges
```

## 6. Examples

### 6a. Season 1 Bracket

```{r Viz_Season1Bracket , eval=TRUE,echo=FALSE,message=FALSE,include=TRUE,warning=FALSE}
library(touRnamentofchampions);library(ggplot2);library(ggbump)

## Set up the data 
  season1bracket <- touRnamentofchampions::results %>% 
    filter(season == 1) %>%
    left_join(touRnamentofchampions::seeds) %>%
    select(!season)

  ## Bump chart data
    bumpdata <- season1bracket %>% select(round,chef,total,seed,coast,challenge,y) %>%
    mutate(challengelong=case_when(round %in% c("Round 1","Round 2","Quarter-final") ~ paste(round,coast,challenge,sep="_"),
                                   round %in% c("Semi-final") & grepl("East",coast) ~"Semi-final_East",
                                   round %in% c("Semi-final") & grepl("West",coast) ~"Semi-final_West",
                                   round %in% c("Final") ~ round)
           ,x=case_when(round=="Round 1" & grepl("East",coast) ~ 85
                        ,round=="Round 1" & grepl("West",coast) ~ 15
                        ,round=="Quarter-final" & grepl("East",coast) ~ 70
                        ,round=="Quarter-final" & grepl("West",coast) ~ 30
                        ,round=="Semi-final" & grepl("East",coast) ~ 60
                        ,round=="Semi-final" & grepl("West",coast) ~ 40
                        ,round=="Final"  ~ 50)
           ,alignment=case_when(grepl("East",coast) & round=="Round 1"~ 0
                                ,grepl("West",coast) & round=="Round 1"~ 1
                                ,TRUE ~ NA)
           #,chefseed=ifelse(grepl("East",coast),paste(chef," (",seed," seed)",sep=""),paste("(",seed," seed) ",chef,sep=""))
    ) %>% 
    group_by(challengelong) %>%
    mutate(sizeofdot=case_when(total < 80 & !(is.na(total)) ~ "75 to 79"
                               ,total >= 80 & total <85 ~ "80 to 84"
                               ,total >= 85 & total <90 ~ "85 to 89"
                               ,total >= 90 & !(is.na(total)) ~ "90+")
           ,cheflabelx=ifelse(grepl("East",coast),x+5,x-5)
           ,winner=ifelse(total==max(total),"Winner","Loser")
           #,w=strwidth(chef,'inches')+0.05,h=strheight(chef,'inches') + 0.5
    ) %>%
    ungroup() %>%
    group_by(sizeofdot) %>%
    mutate(sizeofdot=paste(sizeofdot," (occured ",n()," times)",sep="")) %>%
    ungroup(sizeofdot)
    
  ## Vertical data
      verticaldata <- bumpdata %>%
        select(challengelong,y) %>%
        arrange(challengelong)%>%
        ungroup()
      
      verticaldata <- verticaldata %>%
        mutate(yindic=ifelse(as.numeric(row.names(verticaldata)) %% 2 == 0,"Y2","Y1")) %>%
        pivot_wider(names_from="yindic",values_from="y") %>%
        mutate(Y1=case_when(challengelong == "Final"~6
                            ,challengelong=="Semi-final_West"~11
                            ,challengelong=="Semi-final_East"~11
                            ,TRUE~Y1)) %>%
        left_join(bumpdata %>% select(challengelong,x,coast) %>% distinct()) %>%
        mutate(X2=x)
    
## Bump chart visual
  bumpdata %>%
      ggplot() +
      # add vertical lines to indicate who competed against whom
      geom_segment(data=verticaldata,aes(x=x,xend=X2,y=Y1,yend=Y2),lty=2,color="#57606c") +
      # bump lines
      geom_bump(aes(x=bumpdata$x,y=bumpdata$y,fill=bumpdata$chef),color="#57606c" ) +
      # dots -- and highlight the winners
      geom_point(data=bumpdata,aes(x=x,y=y,color=sizeofdot,size=sizeofdot,shape=sizeofdot),fill="#1170AA" )+
      geom_point(data=bumpdata %>% filter(winner=="Winner"),
                 aes(x=x,y=y),shape=21,color="#57606c",fill="transparent",size=10) +
      annotate("text",x=bumpdata$cheflabelx[bumpdata$round == "Round 1"],y=bumpdata$y[bumpdata$round == "Round 1"]
               ,label=bumpdata$chef[bumpdata$round == "Round 1"]
               ,hjust=bumpdata$alignment[bumpdata$round == "Round 1"]
               ,size=3) +
      annotate("text",x=bumpdata$cheflabelx[bumpdata$round == "Round 1"],y=bumpdata$y[bumpdata$round == "Round 1"]-.7
               ,label=paste(bumpdata$seed[bumpdata$round == "Round 1"],"seed",sep=" ")
               ,hjust=bumpdata$alignment[bumpdata$round == "Round 1"]
               ,size=3) +
      scale_x_continuous(limits=c(-15,115),breaks=c(0,15,30,40,50,60,70,85,100),
                         labels=c("","Round 1","Quarter-\nfinal","Semi-\nfinal",
                                  "Final","Semi-\nfinal","Quarter-\nfinal","Round 1","")) +
      xlab("") + ylab("") +
      theme_minimal() +
      theme(panel.grid = element_blank()
            ,plot.title=element_text(hjust=.5)
            ,plot.subtitle=element_text(hjust=.5)
            ,title=element_text()
            ,axis.text.x=element_text(color="black",size=6)
            ,axis.text.y=element_blank()
            ,legend.text = element_text()
            ,legend.title = element_text() ) +
      scale_color_manual(values=c("#c85200", "#A3ACB9","#a3cce9","#1170AA")
                         ,name="Score",labels=sort(unique(bumpdata$sizeofdot)))+
      scale_shape_manual(values=c(20,16,19,21),name="Score",labels=sort(unique(bumpdata$sizeofdot))) +
      scale_size_manual(values=c(3,4,6,8),name="Score",labels=sort(unique(bumpdata$sizeofdot)) )+
      labs(title="Tournament of Champions I: Winners and scores by round"
           ,subtitle="Winners of each round outlined in black circles"
           ,caption="Data: github.com/celevitz/touRnamentofchampions /// Twitter @carlylevitz") 

```

### 6b. Season 4 Bracket


```{r Viz_Season4Bracket , eval=TRUE,echo=FALSE,message=FALSE,include=TRUE,warning=FALSE}
library(touRnamentofchampions);library(ggplot2);library(ggbump)

## Set up the data 
  season1bracket <- touRnamentofchampions::results %>% 
    filter(season == 4) %>%
    left_join(touRnamentofchampions::seeds) %>%
    select(!season)

  ## Bump chart data
    bumpdata <- season1bracket %>% select(round,chef,total,seed,coast,challenge,y) %>%
    mutate(challengelong=case_when(round %in% c("Round 1","Round 2","Quarter-final") ~ paste(round,coast,challenge,sep="_"),
                                   round %in% c("Semi-final") & grepl("East",coast) ~"Semi-final_East",
                                   round %in% c("Semi-final") & grepl("West",coast) ~"Semi-final_West",
                                   round %in% c("Final") ~ round)
           ,x=case_when(round=="Round 1" & grepl("East",coast) ~ 100
                        ,round=="Round 1" & grepl("West",coast) ~ 0
                        ,round=="Round 2" & grepl("East",coast) ~ 85
                        ,round=="Round 2" & grepl("West",coast) ~ 15
                        ,round=="Quarter-final" & grepl("East",coast) ~ 70
                        ,round=="Quarter-final" & grepl("West",coast) ~ 30
                        ,round=="Semi-final" & grepl("East",coast) ~ 60
                        ,round=="Semi-final" & grepl("West",coast) ~ 40
                        ,round=="Final"  ~ 50)
           ,alignment=case_when(grepl("East",coast) & round=="Round 1"~ 0
                                ,grepl("West",coast) & round=="Round 1"~ 1
                                ,TRUE ~ NA)
           #,chefseed=ifelse(grepl("East",coast),paste(chef," (",seed," seed)",sep=""),paste("(",seed," seed) ",chef,sep=""))
    ) %>% 
    group_by(challengelong) %>%
    mutate(sizeofdot=case_when(total < 80 & !(is.na(total)) ~ "75 to 79"
                               ,total >= 80 & total <85 ~ "80 to 84"
                               ,total >= 85 & total <90 ~ "85 to 89"
                               ,total >= 90 & !(is.na(total)) ~ "90+")
           ,cheflabelx=ifelse(grepl("East",coast),x+5,x-5)
           ,winner=ifelse(total==max(total),"Winner","Loser")
           #,w=strwidth(chef,'inches')+0.05,h=strheight(chef,'inches') + 0.5
    ) %>%
    ungroup() %>%
    group_by(sizeofdot) %>%
    mutate(sizeofdot=paste(sizeofdot," (occured ",n()," times)",sep="")) %>%
    ungroup(sizeofdot)
    
  ## Vertical data
      verticaldata <- bumpdata %>%
        select(challengelong,y) %>%
        arrange(challengelong)%>%
        ungroup()
      
      verticaldata <- verticaldata %>%
        mutate(yindic=ifelse(as.numeric(row.names(verticaldata)) %% 2 == 0,"Y2","Y1")) %>%
        pivot_wider(names_from="yindic",values_from="y") %>%
        mutate(Y1=case_when(challengelong == "Final"~16
                            ,challengelong=="Semi-final_West"~18
                            ,challengelong=="Semi-final_East"~18
                            ,TRUE~Y1)) %>%
        left_join(bumpdata %>% select(challengelong,x,coast) %>% distinct()) %>%
        mutate(X2=x)
    
## Bump chart visual
  bumpdata %>%
      ggplot() +
      # add vertical lines to indicate who competed against whom
      geom_segment(data=verticaldata,aes(x=x,xend=X2,y=Y1,yend=Y2),lty=2,color="#57606c") +
      # bump lines
      geom_bump(aes(x=bumpdata$x,y=bumpdata$y,fill=bumpdata$chef),color="#57606c" ) +
      # dots -- and highlight the winners
      geom_point(data=bumpdata,aes(x=x,y=y,color=sizeofdot,size=sizeofdot,shape=sizeofdot),fill="#1170AA" )+
      geom_point(data=bumpdata %>% filter(winner=="Winner"),
                 aes(x=x,y=y),shape=21,color="#57606c",fill="transparent",size=7) +
      annotate("text",x=bumpdata$cheflabelx[bumpdata$round == "Round 1"],y=bumpdata$y[bumpdata$round == "Round 1"]
               ,label=bumpdata$chef[bumpdata$round == "Round 1"]
               ,hjust=bumpdata$alignment[bumpdata$round == "Round 1"]
               ,size=2.5) +
      annotate("text",x=bumpdata$cheflabelx[bumpdata$round == "Round 1"],y=bumpdata$y[bumpdata$round == "Round 1"]-.7
               ,label=paste(bumpdata$seed[bumpdata$round == "Round 1"],"seed",sep=" ")
               ,hjust=bumpdata$alignment[bumpdata$round == "Round 1"]
               ,size=2) +
      scale_x_continuous(limits=c(-45,140),breaks=c(0,15,30,40,50,60,70,85,100),
                         labels=c("Round 1","Round 2","Quarter-\nfinal","Semi-\nfinal",
                                  "Final","Semi-\nfinal","Quarter-\nfinal","Round 2","Round 1")) +
      xlab("") + ylab("") +
      theme_minimal() +
      theme(panel.grid = element_blank()
            ,plot.title=element_text(hjust=.5)
            ,plot.subtitle=element_text(hjust=.5)
            ,title=element_text()
            ,axis.text.x=element_text(color="black",size=6)
            ,axis.text.y=element_blank()
            ,legend.text = element_text()
            ,legend.title = element_text() ) +
      scale_color_manual(values=c("#c85200", "#A3ACB9","#a3cce9","#1170AA")
                         ,name="Score",labels=sort(unique(bumpdata$sizeofdot)))+
      scale_shape_manual(values=c(20,16,19,21),name="Score",labels=sort(unique(bumpdata$sizeofdot))) +
      scale_size_manual(values=c(2,3,4,5),name="Score",labels=sort(unique(bumpdata$sizeofdot)) )+
      labs(title="Tournament of Champions IV: Winners and scores by round"
           ,subtitle="Winners of each round outlined in black circles"
           ,caption="Data: github.com/celevitz/touRnamentofchampions /// Twitter @carlylevitz") 

```

