---
title: "touRnamentofchampions"

output: github_document

---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE
)
```


## 1. Introduction to touRnamentofchampions

touRnamentofchampions is a collection of data sets detailing events across all seasons of Tournament of Champions. It includes Chef information, randomizer information, challenge results, and judge names.

## 2. Installation

Not yet on CRAN. So please use: *devtools::install.packages("celevitz/touRnamentofchampions")*.
If it's not appearing to be updated, restart your R sessions, install it again, and call it into your library. 
```{r InstallPackage , message=FALSE,warning=FALSE}
devtools::install_github("celevitz/touRnamentofchampions")
```

## 3. News
Issues to work on:

* Data entry for Seasons 2 and 3
* The Randomizer data set has duplicate rows for the finals, one for each coast.
* Complete handedness
* Add gender to seed dataset

## 4. References & Acknowlegements

Data were collected manually while watching each season of Tournament of Champions.

Huge thanks to <https://github.com/doehm> for all his support!

```{r Library set up, message=FALSE, warning=FALSE,echo=FALSE,eval=TRUE}
library(touRnamentofchampions); library(tidyverse)
```

## 5. Datasets

Across datasets, key joining variables include:

* `chef`
* `season`
* `coast`
* `round`
* `episode`

### 5a. Seeds

* `chef`: Chef name (full name)
* `season`: Season number
* `seed`: Seed within their section of the bracket: values of 1 through 8
* `coast`: Are they East or West Coast? And is it sub-bracket A or B?
* `nickname`: Guy Fieri's nickname for the chef
* `handedness`: Whether the chef is righthanded, lefthanded, or ambidextrous


```{r Seeds_table , eval=TRUE} 
seeds 
```

### 5b. Randomizer

* `season`: Season number
* `episode`: Episode number
* `round`: Stage of the tournament: Round 1, Round 2, Quarterfinals, Semifinals, Final
* `challenge`: Variable to help distinguish rounds within the same Coast & Round
* `coast`: The coast depends on how many chefs start the competition. If there are 16 chefs, then the coasts are: East, West. If there are 32 chefs, then the coasts are: East A, East B, West A, West B
* `randomizer1`: First wheel of randomizer
* `randomizer2`: Second wheel of randomizer
* `randomizer3`: Third wheel of randomizer
* `randomizer4`: Fourth wheel of randomizer
* `time`: Length of challenge. Unit is minutes

```{r Randomizer_table , eval=TRUE} 
randomizer 
```

### 5c. Results

* `season`: Season number
* `episode`: Episode number
* `round`: Stage of the tournament: Round 1, Round 2, Quarterfinals, Semifinals, Final}
* `challenge`: Variable to help distinguish rounds within the same Coast & Round
* `coast`: The coast depends on how many chefs start the competition. If there are 16 chefs, then the coasts are: East, West. If there are 32 chefs, then the coasts are: East A, East B, West A, West B
* `y`: Numeric value to help when creating the bracket
* `chef`: Name of chef
* `commentator`:  Who presented their food to the judges: Simon Majumdar or Justin Warner? 
* `order`: When did their food get presented to the judges: Presented 1st or Presented 2nd 
* `score_taste`: Score that chef received for the taste of their dish: values of 0- 50
* `score_randomizer`: Score that chef received for how well they used the Randomizer: values of 0- 30
* `score_presentation`: Score that chef received for the presentation of their dish: values of 0- 20
* `total`: Total score that chef received: between 0 and 100 

```{r Results_table , eval=TRUE} 
results 
```

### 5d. Judges

If a judge scores for multiple rounds across one episode, they will appear more than once in an episode. So unique identifier is `season`-`episode`-`round`.

* `season`: Season number
* `episode`: Episode number
* `round`: Stage of the tournament: Round 1, Round 2, Quarterfinals, Semifinals, Final
* `judge`: Name of guest judge

```{r Judges_table , eval=TRUE} 
judges
```
