---
title: "touRnamentofchampions"

output: github_document
---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE
)
```

## 1. Introduction to touRnamentofchampions

touRnamentofchampions is a collection of data sets detailing events across all seasons of Tournament of Champions. It includes Chef information, randomizer information, challenge results, and judge names.

## 2. Installation

Not yet on CRAN. So please use: *devtools::install.packages("celevitz/touRnamentofchampions")*. If it's not appearing to be updated, restart your R sessions, install it again, and call it into your library.

```{r InstallPackage , message=FALSE,warning=FALSE}
devtools::install_github("celevitz/touRnamentofchampions")
```

## 3. News

Issues to work on:

- Data entry for Season 5 as it airs
- Complete handedness and nicknames

## 4. References & Acknowlegements

Data were collected manually while watching each season of Tournament of Champions.

Huge thanks to <https://github.com/doehm> for all his support!

```{r Library set up, message=FALSE, warning=FALSE,echo=FALSE,eval=TRUE}
library(touRnamentofchampions); library(tidyverse)
```

## 5. Datasets

Across datasets, key joining variables include:

-   `chef`
-   `season`
-   `coast`
-   `region`
-   `round`
-   `episode`
-   `challenge`

See the sections below for additional information on datasets.

<details>
<summary>
<strong>Seeds</strong>
</summary>
### Seeds
The unique identifiers of this dataset are `chef`-`season`.

-   `chef`: Chef name (full name)
-   `season`: Season number
-   `seed`: Seed within their section of the bracket: values of 1 through 8. 
Chefs that played in the play-in but didn't make the final bracket will have 
letters in their seeds.
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If 
there are 16 chefs, then the region is left blank. If there are 32 chefs, then 
the regions are A or B.

```{r Seeds_table , eval=TRUE}
seeds 
```

</details>
<details>
<summary>
<strong>Chefs</strong>
</summary>

### Chefs

The unique identifier of this dataset is `chef`.

-   `chef`: Chef name (full name)
-   `nickname`: Guy Fieri's nickname for the chef
-   `handedness`: Whether the chef is righthanded, lefthanded, or ambidextrous
-   `gender`: male, female, nonbinary

```{r Chefs_table , eval=TRUE}
chefs 
```
</details>
<details>
<summary>
<strong>Randomizer</strong>
</summary>

### Randomizer

The unique identifiers of this dataset are `season`-`episode`-`round`-`challenge`. The reason that `episode` is a unique identifier is because in Season 2, Jet and Antonia tied in all scores and so had a rematch in the Quarter-finals (episodes 6 and 7).

-   `season`: Season number
-   `episode`: Episode number
-   `round`: Stage of the tournament: Play-in, Qualifier semi-final, Qualifier 
final, Round of 32, Round of 16, Quarterfinals, Semifinals, Final
-   `challenge`: Variable to help distinguish rounds within the same Coast & Round
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
-   `randomizer1`: First wheel of randomizer
-   `randomizer2`: Second wheel of randomizer
-   `randomizer3`: Third wheel of randomizer
-   `randomizer4`: Fourth wheel of randomizer
-   `time`: Length of challenge. Unit is minutes
-   `randomizer5`: Fifth wheel of randomizer

```{r Randomizer_table , eval=TRUE}
randomizer 
```

</details>
<details>
<summary>
<strong>Results</strong>
</summary>
### Results

The unique identifiers of this dataset are `season`-`episode`-`round`-`challenge`-`chef`.

-   `season`: Season number
-   `episode`: Episode number
-   `round`: Stage of the tournament: Play-in, Qualifier semi-final, Qualifier final, Round of 32, Round of 16, Quarterfinals, Semifinals, Final
-   `challenge`: Variable to help distinguish rounds within the same Coast & Round
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
-   `y`: Numeric value to help when creating the bracket
-   `chef`: Name of chef
-   `commentator`: Who presented their food to the judges: Simon Majumdar or Justin Warner?
-   `order`: When did their food get presented to the judges: Presented 1st or Presented 2nd
-   `score_taste`: Score that chef received for the taste of their dish: values of 0- 50
-   `score_randomizer`: Score that chef received for how well they used the Randomizer: values of 0- 30
-   `score_presentation`: Score that chef received for the presentation of their dish: values of 0- 20
-   `total`: Total score that chef received: between 0 and 100

```{r Results_table , eval=TRUE}
results 
```

</details>
<details>
<summary>
<strong>Judges</strong>
</summary>

### Judges

The unique identifier is `season`-`episode`-`round`, because occasionally a judge will only judge for one round within an episode.

-   `season`: Season number
-   `episode`: Episode number
-   `judge`: Name of guest judge
-   `round`: Stage of the tournament: Play-in, Qualifier semi-final, Qualifier 
final, Round of 32, Round of 16, Quarterfinals, Semifinals, Final

```{r Judges_table , eval=TRUE}
judges
```
</details>

## 6. Examples

See the sections below for examples of how to use the data

<details>
<summary>
<strong>Gender distribution by season: Seasons 1 through 4</strong>
</summary>
```{r Gender_distr_by_season1to4 , eval=TRUE,echo=FALSE}
seeds %>% left_join(chefs) %>%
  # keep only seasons 1 through 4
  filter(season < 5) %>%
  group_by(season,gender) %>%
  summarise(n=n()) %>%
  pivot_wider(names_from=gender,values_from=n)
```
</details>


