---
title: "touRnamentofchampions"

output: github_document

---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE
)
```


## 1. Introduction to touRnamentofchampions

touRnamentofchampions is a collection of data sets detailing events across all seasons of Tournament of Champions. It includes Chef information, randomizer information, challenge results, and judge names.

## 2. Installation

Not yet on CRAN. So please use: *devtools::install.packages("celevitz/touRnamentofchampions")*.
If it's not appearing to be updated, restart your R sessions, install it again, and call it into your library. 
```{r InstallPackage , message=FALSE,warning=FALSE}
devtools::install_github("celevitz/touRnamentofchampions")
```

## 3. References & Acknowlegements

Data were collected manually while watching each season of Tournament of Champions.

Huge thanks to <https://github.com/doehm> for all his support!

```{r Library set up, message=FALSE, warning=FALSE,echo=FALSE,eval=TRUE}
library(touRnamentofchampions); library(tidyverse)
```

## 4. Overview of datasets

Across datasets, key joining variables include:

* `chef`
* `season`
* `coast`
* `round`
* `episode`
