---
title: "touRnamentofchampions"

output: github_document
---

```{r, include = FALSE, message=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE
)
```

## 1. Introduction to touRnamentofchampions

touRnamentofchampions is a collection of data sets detailing events across all seasons of Tournament of Champions. It includes Chef information, randomizer information, challenge results, and judge names.

## 2. Installation

Not yet on CRAN. So please use: *devtools::install.github("celevitz/touRnamentofchampions")*. If it's not appearing to be updated, restart your R sessions, install it again, and call it into your library.

```{r InstallPackage , message=FALSE,warning=FALSE}
devtools::install_github("celevitz/touRnamentofchampions")
```

## 3. News

Issues to work on:

- Data entry for Season 5 as it airs
- Complete handedness and nicknames

## 4. References & Acknowledgements

Data were collected manually while watching each season of Tournament of Champions.

Huge thanks to <https://github.com/doehm> for all his support!

```{r Library set up, message=FALSE, warning=FALSE,echo=FALSE,eval=TRUE}
library(touRnamentofchampions); library(tidyverse); library(ggbump)
```

## 5. Datasets

Across datasets, key joining variables include:

-   `chef`
-   `season`
-   `coast`
-   `region`
-   `round`
-   `episode`
-   `challenge`

See the sections below for additional information on datasets.

<details>
<summary>
<strong>Seeds</strong>
</summary>
### Seeds
The unique identifiers of this dataset are `chef`-`season`.

-   `chef`: Chef name (full name)
-   `season`: Season number
-   `seed`: Seed within their section of the bracket: values of 1 through 8. Chefs that played in the qualifiers but didn't make the final bracket will have seeds of 8.2, 8.3, or 8.4.
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If 
there are 16 chefs, then the region is left blank. If there are 32 chefs, then 
the regions are A or B.

```{r Seeds_table , eval=TRUE}
seeds 
```

</details>
<details>
<summary>
<strong>Chefs</strong>
</summary>

### Chefs

The unique identifier of this dataset is `chef`.

-   `chef`: Chef name (full name)
-   `nickname`: Guy Fieri's nickname for the chef
-   `handedness`: Whether the chef is righthanded, lefthanded, or ambidextrous
-   `gender`: male, female, nonbinary

```{r Chefs_table , eval=TRUE}
chefs 
```
</details>
<details>
<summary>
<strong>Randomizer</strong>
</summary>

### Randomizer

The unique identifiers of this dataset are `season`-`episode`-`round`-`challenge`. The reason that `episode` is a unique identifier is because in Season 2, Jet and Antonia tied in all scores and so had a rematch in the Quarter-finals (episodes 6 and 7).

-   `season`: Season number
-   `episode`: Episode number
-   `round`: Stage of the tournament: Qualifier semi-final, Qualifier  final, Round of 32, Round of 16, Quarterfinals, Semifinals, Final
-   `challenge`: Variable to help distinguish rounds within the same Coast & Round
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
-   `randomizer1`: First wheel of randomizer
-   `randomizer2`: Second wheel of randomizer
-   `randomizer3`: Third wheel of randomizer
-   `randomizer4`: Fourth wheel of randomizer
-   `time`: Length of challenge. Unit is minutes
-   `randomizer5`: Fifth wheel of randomizer

```{r Randomizer_table , eval=TRUE}
randomizer 
```

</details>
<details>
<summary>
<strong>Randomizer (long form)</strong>
</summary>
### Randomizer (long form)
A dataset containing information about each challenge: protein, vegetables, equipment, style, time. However, it's in "long form" so each challenge shows up multiple times. It categorizes the randomizer ingredients into categories and subcategories.

The unique identifiers of this dataset are `season`-`episode`-`round`-`challenge`-`randomizer`. 

- `season`: Season number
- `episode`: Episode number
- `round`: Stage of the tournament: Qualifier semi-final, Qualifier final, Round of 32, Round of 16, Quarterfinals, Semifinals, Final
- `challenge`: Variable to help distinguish challenges within the same Coast & Round
- `coast`: East or West
- `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
- `time`: Length of challenge. Unit is minutes
- `randomizer`: What wheel was spun (1, 2, 3, 4, or 5)
- `value`: What was the value/item on the randomzier wheel?
- `category`: Categorical variable: protein,produce,equipment,style,wildcard
- `subcategory`: Subcategories for protein (Beef, Fish, Game, Other, Pork, Poultry, Shellfish) and style (Region/country, Style, Theme)

```{r Randomizer_LongForm , eval=TRUE}
randomizerlongform 
```

</details>
<details>
<summary>
<strong>Results</strong>
</summary>
### Results

The unique identifiers of this dataset are `season`-`episode`-`round`-`challenge`-`chef`.

-   `season`: Season number
-   `episode`: Episode number
-   `round`: Stage of the tournament: Qualifier semi-final, Qualifier final, Round of 32, Round of 16, Quarterfinals, Semifinals, Final
-   `challenge`: Variable to help distinguish rounds within the same Coast & Round
-   `coast`: Are they East or West Coast?
-   `region`: The region depends on how many chefs start the competition. If there are 16 chefs, then the region is left blank. If there are 32 chefs, then the regions are A or B.
-   `chef`: Name of chef
-   `commentator`: Who presented their food to the judges: Simon Majumdar or Justin Warner?
-   `order`: When did their food get presented to the judges: Presented 1st or Presented 2nd
-   `score_taste`: Score that chef received for the taste of their dish: values of 0- 50
-   `score_randomizer`: Score that chef received for how well they used the Randomizer: values of 0- 30
-   `score_presentation`: Score that chef received for the presentation of their dish: values of 0- 20
-   `total`: Total score that chef received: between 0 and 100
-   `winner`: Winner, loser, or tie
-   `x`: Numeric X value to help when creating the bracket
-   `y`: Numeric Y value to help when creating the bracket

```{r Results_table , eval=TRUE}
results 
```

</details>
<details>
<summary>
<strong>Judges</strong>
</summary>

### Judges

The unique identifier is `season`-`episode`-`round`, because occasionally a judge will only judge for one round within an episode.

-   `season`: Season number
-   `episode`: Episode number
-   `judge`: Name of guest judge
-   `round`: Stage of the tournament: Qualifier semi-final, Qualifier final, Round of 32, Round of 16, Quarterfinals, Semifinals, Final

```{r Judges_table , eval=TRUE}
judges
```
</details>

## 6. Examples

See the sections below for examples of how to use the data

<details>
<summary>
<strong>Brackets</strong>
</summary>

### Brackets
```{r Brackets , eval=TRUE,echo=FALSE,message=FALSE,include=TRUE,warning=FALSE}
resultsraw <- touRnamentofchampions::results
seeds <- touRnamentofchampions::seeds

results <- resultsraw %>%
  left_join(seeds %>% select(chef,season,seed,coast,region)) %>%
  mutate(sizeofdot=case_when(total < 80 & !(is.na(total)) ~ "71 to 79"
                             ,total >= 80 & total <85 ~ "80 to 84"
                             ,total >= 85 & total <90 ~ "85 to 89"
                             ,total >= 90 & !(is.na(total)) ~ "90+"
                             ,is.na(total) ~ "Not scored")
         ,cheflabelx=ifelse(grepl("East",coast),x+7,x-7)
         ,alignment=case_when(grepl("East",coast) & round=="Round of 32"~ 0
                              ,grepl("West",coast) & round=="Round of 32"~ 1
                              ,TRUE ~ NA)
         ,longchall=paste0(season,episode,round,challenge)
         ) %>%
  group_by(season,sizeofdot) %>%
  mutate(sizeofdot=paste(sizeofdot," (occured ",n()," times)",sep="")) %>%
  ungroup()
  

## Function for bracket:
bracketfunction <- function(seasonnumber,xlimits,xbreaks,xlabs,titlenumber,roundtolabel) {
  ## Bump data
    bumpdata <- results %>% filter(season == seasonnumber)
  
  ## Vertical data
    verticaldata <- bumpdata %>%
      select(longchall,y) %>%
      arrange(longchall)%>%
      ungroup() %>% group_by(longchall) %>%
      mutate(yindic=row_number()
             ,yindic=paste0("Y",as.character(yindic))) %>%
      pivot_wider(names_from="yindic",values_from="y") %>%
      left_join(bumpdata %>% select(longchall,x,coast,winner) %>% distinct()) %>%
      mutate(X2=x)
    
  ## Plot
    ggplot() +
    # add vertical lines to indicate who competed against whom
    geom_segment(data=verticaldata,aes(x=x,xend=X2,y=Y1,yend=Y2),lty=2,color="#57606C") +
    # bump lines
    geom_bump(data=bumpdata,aes(x=x,y=y,fill=chef) ,color="#57606C" ) +
    # dots -- and highlight the winners
    geom_point(data=bumpdata,aes(x=x,y=y,color=sizeofdot,size=sizeofdot,shape=sizeofdot),fill="#1170AA" )+
    geom_point(data=bumpdata %>% filter(winner=="Winner"),
               aes(x=x,y=y),shape=21,color="#57606c",fill="transparent",size=8) +
    scale_shape_manual(values=c(20,16,19,21,4),name="Score legend",labels=sort(unique(bumpdata$sizeofdot))) +
    scale_size_manual(values=c(2,3,4,5,3),name="Score legend",labels=sort(unique(bumpdata$sizeofdot)) ) +
    scale_color_manual(values=c("#c85200", "#A3ACB9","#a3cce9","#1170AA","brown")
                       ,name="Score legend",labels=sort(unique(bumpdata$sizeofdot))) +
    annotate("text",x=bumpdata$cheflabelx[bumpdata$round == roundtolabel],y=bumpdata$y[bumpdata$round == roundtolabel]
             ,label=bumpdata$chef[bumpdata$round == roundtolabel]
             ,hjust=bumpdata$alignment[bumpdata$round == roundtolabel]
             ,size=2) +
    annotate("text",x=bumpdata$cheflabelx[bumpdata$round == roundtolabel],y=bumpdata$y[bumpdata$round == roundtolabel]-.7
             ,label=paste(bumpdata$seed[bumpdata$round == roundtolabel],"seed",sep=" ")
             ,hjust=bumpdata$alignment[bumpdata$round == roundtolabel]
             ,size=2) +
    scale_x_continuous(limits=xlimits,breaks=xbreaks,labels=xlabs) +
    xlab("") + ylab("") +
    theme_minimal() +
    theme(panel.grid = element_blank()
          ,plot.title=element_text(hjust=.5,color="black",size=10,face="bold")
          ,plot.subtitle=element_text(hjust=.5,color="black",size=8)
          ,plot.caption =element_text(color="black",size=10)
          ,axis.text.x=element_text(color="black",size=8)
          ,axis.text.y=element_blank()
          ,legend.text = element_text(color="black",size=7)
          ,legend.title = element_text(color="black",size=8,face="bold")
          ,legend.background = element_rect(fill=NA,color=NA)
          ,legend.position="bottom") +
    labs(title=paste0("Tournament of Champions ",titlenumber,": Winners and final scores by round")
         ,subtitle="Winners of each round outlined in dark gray circles"
         ,caption="Data: github.com/celevitz/touRnamentofchampions /// Twitter @carlylevitz /// Instagram @carly.sue.bear") 

}


## set the inputs for the function
  # season 1
    seasonnumber <- 1
    xlimits <- c(0,100)
    xbreaks <- c(15,30,40,50,60,70,85)
    xlabs <- c("Round of 16","Quarter-\nfinal","Semi-\nfinal",
               "Final","Semi-\nfinal","Quarter-\nfinal","Round of 16")
    titlenumber <- "I"
    roundtolabel <- "Round of 16"
  
    bracketfunction(seasonnumber,xlimits,xbreaks,xlabs,titlenumber,roundtolabel)
    
  # season 2
    seasonnumber <- 2
    xlimits <- c(0,100)
    xbreaks <- c(15,30,40,50,60,70,85)
    xlabs <- c("Round of 16","Quarter-\nfinal","Semi-\nfinal",
               "Final","Semi-\nfinal","Quarter-\nfinal","Round of 16")
    titlenumber <- "II"
    roundtolabel <- "Round of 16"
  
    bracketfunction(seasonnumber,xlimits,xbreaks,xlabs,titlenumber,roundtolabel)
    
  # season 3
    seasonnumber <- 3
    xlimits <- c(-30,130)
    xbreaks <- c(0,15,30,40,50,60,70,85,100)
    xlabs <- c("Round\nof 32","Round\nof 16","Quarter-\nfinal","Semi-\nfinal",
               "Final","Semi-\nfinal","Quarter-\nfinal","Round\nof 16","Round\nof 32")
    titlenumber <- "III"
    roundtolabel <- "Round of 32"  
    
    bracketfunction(seasonnumber,xlimits,xbreaks,xlabs,titlenumber,roundtolabel)
    
  # season 4
    seasonnumber <- 4
    xlimits <- c(-30,130)
    xbreaks <- c(0,15,30,40,50,60,70,85,100)
    xlabs <- c("Round\nof 32","Round\nof 16","Quarter-\nfinal","Semi-\nfinal",
               "Final","Semi-\nfinal","Quarter-\nfinal","Round\nof 16","Round\nof 32")
    titlenumber <- "IV"
    roundtolabel <- "Round of 32"  
    
    bracketfunction(seasonnumber,xlimits,xbreaks,xlabs,titlenumber,roundtolabel)

```

</details>

<details>
<summary>
<strong>Gender distribution by season: Seasons 1 through 4</strong>
</summary>

### Gender distribution by season: Season 1 through 4
```{r Gender_distr_by_season1to4 , eval=TRUE,echo=TRUE}
seeds %>% left_join(chefs) %>%
  # keep only seasons 1 through 4
  filter(season < 5) %>%
  group_by(season,gender) %>%
  summarise(n=n()) %>%
  pivot_wider(names_from=gender,values_from=n) 
```
</details>

<details>
<summary>
<strong>Subcategories of proteins used in battles</strong>
</summary>

### Subcategories of proteins used in battles
```{r Protein_subcategories , eval=TRUE,echo=TRUE}
randomizerlongform %>% 
  group_by(category,subcategory) %>% 
  filter(category %in% c("protein")) %>% 
  summarise(number_of_battles=n())
```
</details>
